---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(flextable)
```

<p align="center">
  <img width="15%" src="http://www.ardata.fr/images/fpeek.svg">
</p>

# fpeek

The goal of fpeek is to help importation of text files from R.

When a text file has unknown characteristics, beeing able 
to have a glance at it helps to get those:

* Use function `file_nlines()` to count lines contained in a file.
* Functions `file_head_show()` and `file_tail_show()` will display 
first and last lines of the file. 

Sometimes encoding is an issue when files generated on a "Windows" machine 
has to be used on a "Linux" machine. Function `file_iconv` will converts 
file content in one encoding to another encoding.


## Installation

```{r 'github-install', eval=FALSE}
devtools::install_github("davidgohel/fpeek")
```

## Example

First download an example file, the file is containing lot of times series 
about power system and has lot of lines.

```{r 'download-file'}
library(fpeek)
power_file <- tempfile(fileext = ".csv")
download.file("https://data.open-power-system-data.org/time_series/2018-03-13/time_series_15min_singleindex.csv", destfile = power_file)
```

### Counting number of lines:


```{r 'nlines-ex'}
file_nlines(power_file)
```

### Show first and last lines of a file

The first lines can be printed in the console with function `file_head_show()`. 

```{r 'head-show-ex'}
file_head_show(power_file, n = 4)
```

`file_head_char()` is also provided to get the first lines as a character vector.

```{r 'head-char-ex'}
file_head_char(power_file, n = 4)
```

The last lines can be printed in the console with function `file_tail_show()`. 

```{r 'tail-show-ex'}
file_tail_show(power_file, n = 4)
```

`file_tail_char()` is also provided to get the last lines as a character vector.

```{r 'tail-char-ex'}
file_tail_char(power_file, n = 4)
```


### File reencoding

```{r 'file-iconv-ex'}
la_cigale <- system.file(package = "fpeek", "datafiles", 
  "cigale-ISO-8859-1.txt")
file_head_char(la_cigale, n = 4)

file_iconv(la_cigale, from = "WINDOWS-1252", to = "UTF-8")
```

Function `file_iconv()` is converting the content of the file 
and print the result in the console. Use `sink()` to create 
a file:

```{r 'file-iconv-sink-ex', eval=FALSE}
file_utf8 <- tempfile()

file_con <- file( file_utf8, open = "wt", encoding = "UTF-8")

sink(file = file_con)
file_iconv(la_cigale, from = "WINDOWS-1252", to = "UTF-8")
sink()

close(file_con)
```

## Comparing speed with other methods:

### `file_nlines`

```{r 'bench-nlines', message=FALSE, warning=FALSE}
library(microbenchmark)

benchmark <- microbenchmark(
  "fpeek::file_nlines" = { file_nlines(power_file) },
  "R.utils::countLines" = { R.utils::countLines(power_file) }, 
  "wc command" = { system(sprintf("wc -l %s", power_file), intern = TRUE) }
  )

benchmark
```

### `file_head_show`

```{r 'bench-head-show', message=FALSE, warning=FALSE, results='hide'}
library(microbenchmark)
benchmark <- microbenchmark(
  "fpeek::file_head_show" = { file_head_show(power_file, n = 5) },
  "base::readLines" = { print(readLines(power_file, n = 5)) }, 
  "head command" = { print(system(sprintf("head -n 5 %s", power_file), intern = TRUE)) }
  )
```

```{r}
benchmark
```

