---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(flextable)
```

[![Travis-CI Build Status](https://travis-ci.org/davidgohel/fpeek.svg?branch=master)](https://travis-ci.org/davidgohel/fpeek)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/davidgohel/fpeek?branch=master&svg=true)](https://ci.appveyor.com/project/davidgohel/fpeek)
[![version](http://www.r-pkg.org/badges/version/fpeek)](https://CRAN.R-project.org/package=fpeek)
![cranlogs](http://cranlogs.r-pkg.org./badges/fpeek)
[![Coverage status](https://codecov.io/gh/davidgohel/fpeek/branch/master/graph/badge.svg)](https://codecov.io/github/davidgohel/fpeek?branch=master)


<p align="center">
  <img width="15%" src="http://www.ardata.fr/images/fpeek.svg">
</p>

## fpeek

The goal of fpeek is to help importation of text files from R.

When a text file has unknown characteristics, beeing able 
to have a glance at it helps to get those:

* Use function `peek_count_lines()` to count lines contained in a file.
* Functions `peek_head()` and `peek_tail()` will display 
first and last lines of the file. 

Sometimes encoding is an issue when files generated on a "Windows" machine 
has to be used on a "Linux" machine. Function `peek_iconv` will converts 
file content in one encoding to another encoding.

Performances are good with large files because operations are made 
without reading the entire file before starting. 


## Installation

```{r 'github-install', eval=FALSE}
devtools::install_github("davidgohel/fpeek")
```

## Examples

First download an example file, the file is containing lot of times series 
about power system and has lot of lines.

```{r 'download-file'}
library(fpeek)
power_file <- "singleindex.csv"
if( !file.exists(power_file))
  download.file(
    "https://data.open-power-system-data.org/time_series/2018-03-13/time_series_15min_singleindex.csv", 
    destfile = power_file)
```

### Counting number of lines:

Using `wc -l`

```{r 'wcl-ex'}
system(sprintf("wc -l %s", power_file), intern = TRUE)
```

Using `peek_count_lines()`

```{r 'nlines-ex'}
peek_count_lines(power_file)
```

### Show first and last lines of a file

The first lines can be printed in the console with function `peek_head()`. 

```{r 'head-show-ex'}
peek_head(power_file, n = 4)
```

The last lines can be printed in the console with function `peek_tail()`. 

```{r 'tail-show-ex'}
peek_tail(power_file, n = 4)
```

### File reencoding

```{r 'file-iconv-ex'}
la_cigale <- system.file(package = "fpeek", "datafiles", 
  "cigfou-ISO-8859-1.txt")
peek_head(path = la_cigale, n = 4, intern = TRUE)

peek_iconv(la_cigale, from = "ISO-8859-1", to = "UTF-8")
```

Function `peek_iconv()` is converting the content of the file 
and print the result in the console. Use `peek_iconv(newfile = "...")` to create 
a new reencoded file:

```{r 'file-iconv-sink-ex'}
file_utf8 <- tempfile()
peek_iconv(path = la_cigale, newfile = file_utf8, 
  from = "ISO-8859-1", to = "UTF-8")
```



